\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-06 trunk/europa-basics.tex
re-working EUROPA's plan representation. checkpoint before lunch.

\eu is a versatile planning framework which can be used in a range of
problem-solving methods.

\paragraph \texttt{Constraint Satisfaction}: A canonical problem in
dealing with constraints is the $N$-Queens problem in which chess queens
must be placed on an  $N$x$N$ chessboard so no queens attack the other. Fig.
\ref{fig:nqueens-1} shows an example of a random positioning of queens
on a  $N$x$N$ chessboard. Queens in violation of the non-attack constraint
are highlighted in red.

@@

A constraint programming problem consists of a set of variables $V=
{x_1,..,x_N}$, where each variable takes values from a domain
$d_1,..,d_N$; in this chapter we will deal only with discrete and
finite domains. Given a defined conjunctive set of constraints on the
variables: $C=\{c_1(x_1,..,x_N), ..., c_K(x_1,..,x_N)\}$, the objective
is to find one or more value assignments to $V$ where all constraints
are satisfied.

To solve a problem, CP techniques use logical inference to perform two
important operations:

@@

CP is normally implemented as part of a programming language and
constraints are normally represented as objects \cite{puget95}
\comment{citation}. Any constraint can be introduced and, as long as the
bounds propagation and domain reduction protocols specified by the host
CP system are enforced, it will be indistinguishable from any other
``primitive'' CP constraint, such as $\leq$ or $\geq$. In theory,
solving a CSP problem is NP-Hard \cite{ghallab04}, but often very
efficient in practice using a number of algorithms and techniques that
are well understood.

\comment{TODO: talk about arc consistency algorithms, computational
  complexity and solvers.}

@@

\eu is an implementation of the CAIP framework.

Variables

Values that need to be represented to describe the problem domain and over which we may want to specify constraints.

Objects

The things we wish to describe and refer to in a domain are considered
Objects. As in the case with object-oriented analysis and design, one
can seek out the nouns in any domain description to find likely
objects. In a Mars rover example, we might consider the satellite, the
camera, and the attitude controller (which orients the robot for
taking pictures) to be objects. Objects have state and behavior. For
example, a camera can be:

    off,
    ready, or
    taking a picture.

An attitude controller can be:

    pointing at a position, or
    turning from one position to another.

    An object is an instance of a class. In \eu we model using the
    abstraction of a class to speak about all instances having certain
    properties of state and behavior. In order to describe such state
    and behavior we build on the formalism of first order logic.

Temporally Qualified Predicates and Actions

A predicate defines a relation between objects and properties. In \eu,
we define such relations between variables whose domains are sets of
objects and sets of properties. For example, we might use a predicate
\texttt{Pointing($a,p$)} to indicate that a rover's attitude controller
$a$ is pointing at position $p$. Note that $a$ is a variable which may
have a number of possible values in a problem with multiple
satellites. Similarly, $p$ is a variable whose values are the set of
possible positions. In a grounded plan, single values will be
specified for each variable as we saw in the case of a solved
Constraint Satisfaction Problem.

In general, it is not sufficient to state that a predicate is true
without giving it some temporal extent over which it holds. Predicates
that are always true can be thought to hold from the beginning to the
end of time. However, in practice, the temporal extent of interest
must be defined with timepoints to represent its start and end. So we
might want to write \texttt{Pointing($a,s,e,p$)} to indicate that the
attitude controller a is pointing at position $p$ from time $s$ to
time $e$. In fact, this pattern of using such predicates to describe
both state and behavior of objects is so prevalent in \eu that we have
introduced a special construct called a Token which has the built in
variables to indicate the object to which the statement principally
applies and the timepoints over which it holds. In \eu, all predicate
instances are Tokens.

Tokens

A Token is an instance of a predicate or an action and is defined over
a temporal extent. Every token has five built-in variables:

start: The beginning of the temporal extent over which the predicate
is defined.  end: The end of the temporal extent over which the
predicate is defined.  duration: The constraints start + duration =
end is enforced automatically.  object: made. The set of objects to
which a token might apply. In a grounded plan each Token applies to a
specific Object, reflecting the intuition that we are using Tokens to
describe some aspect of an Object (i.e. its state or behavior) in
time. However, in a partial plan, the commitment to a specific object
may not yet have been made.

state: Tokens can be ACTIVE, INACTIVE, MERGED, or REJECTED. The state
variable captures the token's current state and its reachable states
through further restriction. See the Token State Model discussion
below for details.

Objects (continued)

@@

The state of a token is embodied with a 5th and final built-in
variable referred to as the state variable. The planner states are
values in the domain of this variable {MERGED, ACTIVE, REJECTED}. The
INACTIVE state is captured by the variable being unbound. If a value
is removed from the domain, that state will not be reachable. For
example, when a slave is created via a rule, the REJECTED value is
removed from its state variable.  Summary

This page has presented the main elements of the \eu planning
approach. They are:

Temporally scoped predicates and actions (a.k.a. Tokens) to represent
states and actions in time. Note that tokens do not discriminate
between state and action.

Constraints to describe relationships among tokens. This provides an
expressive method of describing interactions among tokens in the
context of Temporal Planning.

Timelines as a concise abstraction to express the evolution of state
and behavior for a system component. It provides semantics of mutual
exclusion sequencing in time. Other core abstractions are available in
\eu for handling metric resources.

Domain rules to describe internal and external relationships on and
between tokens respectively. Rules are applied to active tokens
(master tokens) referred to as masters and typically produce inactive
tokens (slaves) which are a key vehicle through which the planning
process evolves. The rule structure in \eu differs from the more
restrictive commitment to preconditions and effects in classical
planning which prohibits durative actions and disembodied effects
(i.e. effects which may not occur until some temporal distance after
the end of the action).

Token States which are the basis of planning operations on a partial
plan.

>>>>

At the core of the \rx framework, is the deliberation engine, \eu,
which has a rich legacy from NASA missions. \eu is a versatile
Constraint-based temporal planner which continues to be deployed on a
diverse set of applications. We motivate this section with some
problem domains this planner can handle and then delve in substantial
detail.

% \eu is a versatile planning framework.  let's start by looking at some examples of the main problem types that it can tackle.

\paragraph \textit{Constraint Satisfaction}: A canonical problem in
dealing with constraints is the $N$-Queens problem in which chess queens
must be placed on an  $N$x$N$ chessboard so no queens attack the other. Fig.
\ref{fig:nqueens-1} shows an example of a random positioning of queens
on a  $N$x$N$ chessboard. Queens in violation of the non-attack constraint
are highlighted in red.

@@

A constraint programming problem consists of a set of variables $V=
{x_1,..,x_N}$, where each variable takes values from a domain
$d_1,..,d_N$; in this chapter we will deal only with discrete and
finite domains. Given a defined conjunctive set of constraints on the
variables: $C=\{c_1(x_1,..,x_N), ..., c_K(x_1,..,x_N)\}$, the
objective is to find one or more value assignments to $V$ where all
constraints are satisfied. To solve a problem, CP techniques use
logical inference to perform Constraint Propagation, which consists of
two operations:

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to be arc consistent if for every
value in $x$'s domain, there is a value in $y$'s domain that is
consistent with all the constraints that connect $x$ and $y$.  A naive
approach to ensuring arc consistency consists of cycling through all
the variable pairs and performing constraint propagation until there
are no variable domain changes. The AC-3 algorithm \cite{mackworth77},
a popular implementation, improves over the naive approach by ignoring
constraints whose variables were not affected since the last
iteration.

Arc Consistency and other, more stringent consistency definitions can be
used to efficiently prove that a CP problem is unsatisfiable. However,
they generally cannot prove that a CP problem is satisfiable and
provide specific variable assignments that constitute solutions. For
that, search algorithms like global backtracking \comment{citation
  needed} or local search are used; these use consistency as an
efficient way to prune the search space \comment {citation CP
  Handbook}. In theory, solving a CP problem is NP-Hard
\cite{ghallab04}, but often very efficient in practice using a number
of consistency and search algorithms that are well understood.

CP is usually implemented as part of a programming language and
constraints are usually represented as objects \cite{puget95}. Any
constraint that the user is able to reason about can be introduced into the system and, as long as the constraint
propagation protocols specified by the host CP system are enforced, it
will be indistinguishable from any other "primitive'' CP constraint,
such as $\leq$ or $\geq$.

@@

\eu is an implementation of the CAIP framework. 
\comment{once the CAIP section is finished, this section will need an iteration to provide a few pointers back to it}

\textit{Variables} Values that need to be represented to describe the
problem domain and over which we may want to specify constraints. In the Shopping Agent example, the times at which the agent needs to leave or be back, or executes a purchase, are all instances where variables representing time would be used.

\textit{Objects} The things we wish to describe and refer to in a
domain are considered Objects. As in the case with object-oriented
analysis and design, one can seek out the nouns in any domain
description to find likely objects. In the Shopping Agent example, we night consider the Agent itself, Products and Locations all to be objects. Objects have state and behavior. For example, a Shopping Agent can have:
\begin{enumerate}
	\item State: it's location, a bag that contains the products it has already purchased (the bag can in turn be another object), etc.
	\item Behavior: go to a location to look for a product, perform a purchase, return home, etc. 
\end{enumerate}

As in object-oriented design, objects that have similar state and behavior can be described generically in terms of object types or classes. \eu allows the definition of object classes in the same way that it is done in popular object-oriented programming languages. However, to describe state and behavior for the purposes of planning we need a different mechanism, for that we build on the formalism of first order logic as explained below.

\textit{Tokens}

In first order logic, a predicate defines a relation between objects and properties \comment{citation?}. 
In \eu, we define such relations between variables whose domains are sets of
objects and sets of properties to describe state and behavior. For example, we might use a predicate
\texttt{At($a$,$l$)} to indicate that agent
$a$ is at location $l$, or a predicate \texttt{Buy($a$,$p$}) to indicate that agent $a$ is taking acting to buy product $p$. Note that $a$ is a variable which may
have a number of possible values in a problem with multiple
agents. Similarly, $l$ and $p$ are variables whose values are the set of
possible positions and products respectively. \eu can be used to create partial plans, where the domains of variables $a$ and $l$ can have more than one possible value in them, or grounded plans, where single values will be
specified for each variable as we saw in the case of the N-queen's problem.

In general, to come up with an executable plan it is not sufficient to state predicates that describe required state or behavior
without also specifying some temporal extent over which each of those predicates hold. Predicates
that are always true can be thought to hold from the beginning to the
end of time. However, in practice, the temporal extent of interest
must be defined with timepoints to represent its start and end. So we
might want to write \texttt{At($a$,$s$,$e$,$l$)} to indicate that the
agent $a$ is at location $p$ from time $s$ to
time $e$. In fact, this pattern of using such predicates to describe
both state and behavior of objects is so prevalent in \eu that we have
introduced a special construct called a Token which has the built in
variables to indicate the object to which the statement principally
applies and the timepoints over which it holds. In \eu, all predicate
instances are Tokens. Also, in the same way that objects are described generically by classes, in \eu Tokens can be described generically by Token Types.

A Token is an instance of a predicate that represents and object's state or behavior and is defined over
a temporal extent. Every token has five built-in variables:

\begin{enumerate}
    \item \textit {start}: The beginning of the temporal extent over which the predicate
is defined.  
    \item \textit {end}: The end of the temporal extent over which the
predicate is defined.  
    \item \textit {duration}: The constraint \textit{start} + \textit{duration} =
\textit{end} is enforced automatically.  
    \item \textit{object}: The set of objects to
which a token might apply. In a grounded plan each Token applies to a
specific Object, reflecting the intuition that we are using Tokens to
describe some aspect of an Object (i.e. its state or behavior) in
time. However, in a partial plan, the commitment to a specific object
may not yet have been made.
    \item \textit{state}: Tokens can be ACTIVE, INACTIVE, MERGED, or REJECTED. The state
variable captures the token's current state and its reachable states
through further restriction.  This is \eu's mechanism to support CAIP's planning approach, see the Token State Model discussion
below for details.
\end{enumerate}

\comment {resume cleanup here}

\textit{Objects (continued)}

@@

The state of a token is embodied with a 5th and final built-in
variable referred to as the state variable. The planner states are
values in the domain of this variable {MERGED, ACTIVE, REJECTED}. The
INACTIVE state is captured by the variable being unbound. If a value
is removed from the domain, that state will not be reachable. For
example, when a slave is created via a rule, the REJECTED value is
removed from its state variable.

\textit{Summary}

This has presented the main elements of the EUROPA planning approach. They are:

\item Temporally scoped predicates and actions (a.k.a. Tokens) to represent
states and actions in time. Note that tokens do not discriminate
between state and action.

\item Token States which are the basis of planning operations on a partial
plan.

\item Constraints to describe relationships among tokens. This provides an
expressive method of describing interactions among tokens in the
context of Temporal Planning.

\item Timelines as a concise abstraction to express the evolution of state
and behavior for a system component. It provides semantics of mutual
exclusion sequencing in time. Other core abstractions are available in
\eu for handling metric resources.

\item Domain rules to describe internal and external relationships on and
between tokens respectively. Rules are applied to active tokens
(master tokens) referred to as masters and typically produce inactive
tokens (slaves) which are a key vehicle through which the planning
process evolves. The rule structure in \eu differs from the more
restrictive commitment to preconditions and effects in classical
planning which prohibits durative actions and disembodied effects
(i.e. effects which may not occur until some temporal distance after
the end of the action).
